Evolution of pre-copulatory and post-copulatory inbreeding behaviour: a computational

### 2016-09-16 ms_pre_post.Rmd
Correct standard deviation to standard error in reference to plot shading

To analyse our model, we track mean values of alleles underlying tendency for polyandry ($P_{a}$), inbreeding mating strategy ($M_{a}$), and inbreeding fertilisation strategy ($F_{a}$) in a focal population. We present evolutionary dynamics of these mean allele values across different parameter combinations to test our four focal hypotheses (see below). We simulate up to eight possible permutations of values of costs of tendency for polyandry ($c_{P}$), inbreeding mating strategy ($c_{M}$), and inbreeding fertilisation strategy ($c_{F}$) equal to either $0$ or $0.02$ [@Duthie observed $0.02$ to be a strong cost of tendency for polyandry] for values of $\beta = 3$ ($8$ total combinations). In all simulations, we record mean values of $P_{a}$, $M_{a}$, $F_{a}$ in each generation. Each combination of parameter values simulated is replicated $40$ times, and grand mean values and standard deviation of means is calculated in each generation across replicate simulations. These analyses allow us to infer how allele values change over generations in response to costs, but also in response to the changing values of other alleles and therefore potentially interacting phenotypes (e.g., selection on inbreeding fertilisation strategy might be weak if polyandry is rare). For all replicates, we set the maximum number of generations to 10000, which exploratory simulations and previous modelling [@Duthie] showed to be a sufficient number of generations for inferring long-term dynamics of mean allele values and therefore selection on phenotypes. For all replicates, we also set the number of immigrants per generation to $\rho=5$, which produced a range of kin and non-kin in each generation allowing females could express inbreeding strategies. Females always produced $n=8$ offspring, which was sufficient to keep populations consistently near carrying capacities and avoid population extinction. Carrying capacities for females ($K_{f}$) and males ($K_{m}$) were both set to 100 because previous modelling has shown that populations of this size are small enough that mate encounters between kin occur with sufficient frequency for selection on inbreeding strategy, but populations are also not so small that selection is typically overwhelmed by genetic drift (Duthie and Reid *in press*).

@@

We present our results by showing mean allele value changes over generations. Figures 1-5 show mean allele values of tendency for polyandry ($P_{a}$; red lines and shading), inbreeding mating strategy ($M_{a}$; blue lines and shading), and inbreeding fertilisation strategy ($F_{a}$; black lines and shading) over 10000 generations as calculated across all individuals (female and male) within a focal population over 40 replicate simulations with identical starting conditions. Solid lines represent the grand mean of replicate mean allele values, and shading around lines shows standard deviations of mean allele values around the grand mean. Panels within figures show different cost combinations for tendency for polyandry ($c_{P}$), inbreeding mating strategy ($c_{M}$), and inbreeding fertilisation strategy ($c_{F}$). Figures have fewer than eight panels if some costs are not relevant because allele values are fixed, and fixed allele values are shown as horizontal lines where line height is set to the fixed allele value. Supporting Information S1 presents correlations of mean allele values across replicate simulations, which were generally weak; we found no evidence suggesting strong evolutionary divergence among replicates with identical starting conditions (e.g., as might be the case if populations stochastically evolved toward inbreeding avoidance through mate choice or fertilisation, but not both), meaning that inspecting grand means of replicate allele values accurately reflects expected evolutionary dynamics over generations.

@@

Inbred offspring are typically less viable than outbred offspring, and when such inbreeding depression is strong, it is predicted to cause selection for inbreeding avoidance [Duthie and Reid *in press*; @Parker1979; @Parker2006; @Pusey1996; @Kokko2006]. In animals, inbreeding avoidance can be realised through both pre-copulatory and post-copulatory mechanisms [e.g., @Tuni2013; @Firman2015; @Daniel2015; @Fischer2015]. However, the conditions under which inbreeding avoidance is predicted to evolve via pre-copulatory, post-copulatory, or both pre- and post-copulatory mechanisms has not been modelled. There is therefore no framework for predicting what mechanisms of inbreeding avoidance are most likely to evolve in empirical populations where risk of fitness loss caused by inbreeding is high. Importantly, evolution of post-copulatory mechanisms of inbreeding avoidance requires females to mate multiply, meaning that post-copulatory inbreeding avoidance cannot occur unless polyandry has evolved in a population. Accordingly, we modelled evolution of three separate traits affecting polyandry and inbreeding avoidance, including tendency to engage in polyandry, inbreeding mating strategy (pre-copulatory), and inbreeding fertilisation strategy (post-copulatory). Using *in silico* experiments, we tested four specific hypothesis to predict evolution of tendency for polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy given diverse costs and evolutionary constraints. We found that (1) selection for polyandry was greatly strengthened given concurrent evolution of inbreeding fertilisation strategy, (2) selection was strong for low cost, and weak but not negligible for high cost, mechanisms of inbreeding avoidance, (3) constraints on evolution of pre-copulatory inbreeding avoidance greatly strengthened selection for post-copulatory inbreeding avoidance, but not vice versa, and (4) pre-existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of post-copulatory inbreeding avoidance, but not vice versa. Our results therefore provide a novel framework for approaching empirical studies of inbreeding avoidance, predicting evolution of polyandry and inbreeding avoidance through pre- and post-copulatory mechanisms.

@@

The opportunity to adjust inbreeding has been widely considered as a driver of adaptive evolution of polyandry [@Tregenza2002; @Foerster2003; @Akcay2007; @Varian-Ramos2012; @Kingma2013; @Lehtonen2015; @Reid2014]. Recently, @Duthie found the conditions under which selection for polyandry to avoid inbreeding is predicted to be highly restrictive, requiring strong constraints on the availability of initial compared to additional mates, or the conditional expression of polyandry such that females only engaged in polyandry when their initial choice of mate was of low quality. However, @Duthie assumed that fertilisation of offspring was unbiased with respect to mate kinship, meaning that females could not express choice after mating had occurred. In contrast, we found that when females could express choice after mating through biased fertilisation, selection for polyandry was greatly strengthened (Figure 1A), especially when constraints restricted evolution of inbreeding avoidance through mate choice (Figure 3). The proposition that polyandry might facilitate female choice among males of varying compatability is not new [e.g., @Zeh1997], but the existence of selection for post-copulatory inbreeding avoidance causing feedback to select for polyandry has not been formally modelled. Our model therefore has widespread implications for empirical studies of evolution of polyandry.

We predict evolution of polyandry in populations where inbreeding depression is severe and evolution of inbreeding avoidance through post-copulatory mechanisms is not constrained. Post-copulatory mechanisms of inbreeding avoidance have been observed under these conditions in experimental systems across diverse taxa [e.g. @Pizzari2004; @Firman2008; @Bretman2009; @Gasparini2011; @Tuni2013; @Firman2015]. For example, @Pizzari2004 find that the number of sperm arriving to eggs is lower for full siblings of female red jungefowl (*Gallus gallus*) than it is for non-relatives, even after controlling for effects of order of oviposition and social familiarity. Similarly, Firman and Simmons [-@Firman2008; -@Firman2015] have shown evidence of post-copulatory inbreeding avoidance in house mice, finding fertilisation to be biased toward unrelated males independent of the order in which mating occurs [@Firman2008], perhaps due to post-copulatory mechanisms expressed through the secretion of gametic proteins originating in female's eggs [@Firman2015]. Interestingly, female house mice have also been shown to express pre-copulatory inbreeding avoidance  [@Potts1991; @Roberts2003]. Our results suggest that post-copulatory inbreeding avoidance is unlikely to evolve when females can already avoid inbreeding through pre-copulatory mechanisms (Figure 5), meaning that evolution of post-copulatory inbreeding avoidance could reasonably be hypothesised to have preceded evolution of pre-copulatory inbreeding avoidance in mice.

Specific mechanisms of post-copulatory inbreeding avoidance are difficult to identify, but recent experimental studies have shown how females can successfully bias fetilisation after mating has occurred in both vertebrates [e.g., @Gasparini2011] and invertebrates [e.g., @Tuni2013]. For example, @Gasparini2011 artificially inseminated female guppies (*Poecilia reticulata*) with equal quantities of sperm from full-siblings and unrelated males, finding that fertilisation success was higher in non-relatives because the velocity of full siblings' sperm was reduced by females' ovarian fluids. Biased fertilisation toward non-relatives might also be accomplished through selective storage of sperm. For example, in black field crickets (*Teleogryllus commodus*), females attempt to remove the spermataphores of unwanted male mates after copulation, while male mates guard females to prevent spermatophore removal [@Bussiere2006]. Even if females are unable to remove spermatophores of unwanted mates, @Tuni2013 found evidence that female *T. commodus* are capable of controlling which males' sperm is transferred to their spermatheca, most likely through physical control of the spermathical duct, and reduce sperm storage when being guarded by a related male. Our results suggest that evolution of such post-copulatory inbreeding avoidance might increase selection for polyandry, especially if pre-copulatoyr inbreeding avoidance is costly (Figure 1) or constrained (Figure 3).

@@

Our results show that direct negative selection on phenotypes that increase female promiscuity or choosiness with respect to kinship decrease the strength of selection for tendency for polyandry and inbreeding avoidance (Figure 1). Importantly, the cost of tendency for polyandry only decreased female fitness given tendency for polyandry phenotype values greater than zero ($P_{p}>0$), but the threshold nature of the tendency for polyandry phenotype only resulted in selection against positive tendency for polyandry allele values ($P_{a}$) when the sum over a focal female's $20$ $P_{a}$ alleles was greater than zero because summations less than zero all resulted in equivalent monandrous phenotypes ($P_{p}=0$). Hence, even when tendency for polyandry was costly, selection for more negative $P_{a}$ values weakened as the frequency of negative $P_{a}$ values increased because each individual $P_{a}$ then became less influential in causing females to be monandrous [@Duthie]. This general property of threshold traits causes maladaptive phenotypes to frequently reappear within populations due to the mutation and recombination of continuous genetic values underlying discrete phenotypes [@Roff1996; @Roff1998]. Assuming that there is continuous genetic variation underlying threshold polyandry, polyandrous phenotypes are therefore sometimes predicted even when costly [@Duthie], potentially creating the possibility for selection on inbreeding fertilisation strategy when polyandry is expressed. This is most evident in Figure 1C, where mean allele values for polyandry are negative (red line), but strong post-copulatory inbreeding avoidance is still predicted (grey line). Consequently, costly polyandry does not necessarily preclude evolution of adaptive post-copulatory inbreeding avoidance in populations.

A low relative cost for one inbreeding strategy (e.g., pre-copulatory inbreeding avoidance) did not preclude a more costly complementary inbreeding strategy from persisting in a focal population (Figure 1), meaning that evolution of both pre- and post-copulatory inbreeding avoidance might be expected in populations despite cost asymmetry. However, existence of costs strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003; @Reid2015]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which we suggest are also relevant costs for polyandry, and include costs of elevated risks associated with (1) predation or (2) disease transmission, and costs incurred through loss of time (3) or depletion of energy (4). In our model, we interpreted a cost of tendency for polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of inbreeding mating strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of inbreeding fertilisation strategy as depletion of energy [e.g., @Tuni2013]. While these costs are biologically realistic, different costs will be relevant for different populations and be realised in different ways. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry (as we model). Future models will therefore need to carefully consider the magnitudes and realisation of direct costs to accurately predict evolution of inbreeding strategies in populations.

@@

@@

![Mean allele values underlying tendency for polyandry (red lines and shading), inbreeding mating strategy (i.e., pre-copulatory choice; blue lines and shading), and inbreeding fertilisation strategy (i.e., post-copulatory choice; black lines and shading), as calculated across all individuals within a population over 10000 generations given strong inbreeding depression. Solid lines represent grand means over 40 replicate populations, and shading around solid lines show standard deviations of allele values around grand means. Negative mean allele values reflect strategies of inbreeding avoidance or tendency for monandry, and positive values reflect strategies of inbreeding preference or tendency for polyandry. Panels A-H show simulations for different combinations of costs for tendency for polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy. Dotted horizontal black lines in panels show values of zero on the y-axis.](Initial_Results/figures/evo_combs02_B3.png)

![Mean allele values underlying tendency for polyandry (red lines and shading) and inbreeding mating strategy (i.e., pre-copulatory choice; blue lines and shading), as calculated across all individuals within a population over 10000 generations given strong inbreeding depression when allele values underlying inbreeding fertilisation strategy (i.e., post-copulatory choice) are fixed for random fertilisation among mates (black horizontal lines where y-axes equal zero). Solid lines represent grand means over 40 replicate populations, and shading around solid lines show standard deviations of allele values around grand means. Negative mean allele values reflect strategies of inbreeding avoidance or tendency for monandry, and positive values reflect strategies of inbreeding preference or tendency for polyandry. Panels A-D show simulations for different combinations of costs for tendency for polyandry and inbreeding mating strategy.](Initial_Results/figures/evo_combs02_B3_ko_F.png)

![Mean allele values underlying tendency for polyandry (red lines and shading) and inbreeding fertilisation strategy (i.e., post-copulatory choice; black lines and shading), as calculated across all individuals within a population over 10000 generations given strong inbreeding depression when allele values underlying inbreeding mating strategy (i.e., pre-copulatory choice) are fixed for random mating (blue horizontal lines where y-axes equal zero). Solid lines represent grand means over 40 replicate populations, and shading around solid lines show standard deviations of allele values around grand means. Negative mean allele values reflect strategies of inbreeding avoidance or tendency for monandry, and positive values reflect strategies of inbreeding preference or tendency for polyandry. Panels A-D show simulations for different combinations of costs for tendency for polyandry and inbreeding mating strategy.](Initial_Results/figures/evo_combs02_B3_ko_M-0.png)

![Mean allele values underlying inbreeding mating strategy (i.e., pre-copulatory choice; blue lines and shading), as calculated across all individuals within a population over 10000 generations given strong inbreeding depression and the pre-existence of polyandry to avoid inbreeding (tendency for polyandry allele values equal 1, as indicated by horizontal red lines; inbreeding fertilisation strategy values equal -5, as indicated by horizontal black lines). Solid blue lines represent grand means over 40 replicate populations, and blue shading around solid lines shows standard deviations of mean inbreeding mating strategy allele values around grand means. Negative mean allele values reflect a strategy of inbreeding avoidance, and positive values reflect a strategy of inbreeding preference. Panels A and B show replicate simulations in which inbreeding mating strategy is not costly and has a cost of 0.02, respectively. Dotted horizontal black lines in panels show values of zero on the y-axis.](Initial_Results/figures/evo_combs02_B3_ko_FP-5.png)

![Mean allele values underlying tendency for polyandry (red lines and shading) and inbreeding fertilisation strategy (i.e., post-copulatory choice; black lines and shading), as calculated across all individuals within a population over 10000 generations given strong inbreeding depression and the pre-existence of inbreeding avoidance in mate choice (inbreeding mating strategy values equal -5, as indicated by horizontal blue lines). Solid lines represent grand means over 40 replicate populations, and shading around solid lines show standard deviations of allele values around grand means. Negative mean allele values reflect strategies of inbreeding avoidance or tendency for monandry, and positive values reflect strategies of inbreeding preference or tendency for polyandry. Panels A-D show simulations for different combinations of costs for tendency for polyandry and inbreeding mating strategy. Dotted horizontal black lines in panels show values of zero on they-axis. ](Initial_Results/figures/evo_combs02_B3_ko_M-5.png)

>>>>

To analyse our model, we track mean values of alleles underlying tendency for polyandry ($P_{a}$), inbreeding mating strategy ($M_{a}$), and inbreeding fertilisation strategy ($F_{a}$) in a focal population. We present evolutionary dynamics of these mean allele values across different parameter combinations to test our four focal hypotheses (see below). We simulate up to eight possible permutations of values of costs of tendency for polyandry ($c_{P}$), inbreeding mating strategy ($c_{M}$), and inbreeding fertilisation strategy ($c_{F}$) equal to either $0$ or $0.02$ [@Duthie observed $0.02$ to be a strong cost of tendency for polyandry] for values of $\beta = 3$ ($8$ total combinations). In all simulations, we record mean values of $P_{a}$, $M_{a}$, $F_{a}$ in each generation. Each combination of parameter values simulated is replicated $40$ times, and grand mean values and standard error of means is calculated in each generation across replicate simulations. These analyses allow us to infer how allele values change over generations in response to costs, but also in response to the changing values of other alleles and therefore potentially interacting phenotypes (e.g., selection on inbreeding fertilisation strategy might be weak if polyandry is rare). For all replicates, we set the maximum number of generations to 10000, which exploratory simulations and previous modelling [@Duthie] showed to be a sufficient number of generations for inferring long-term dynamics of mean allele values and therefore selection on phenotypes. For all replicates, we also set the number of immigrants per generation to $\rho=5$, which produced a range of kin and non-kin in each generation allowing females could express inbreeding strategies. Females always produced $n=8$ offspring, which was sufficient to keep populations consistently near carrying capacities and avoid population extinction. Carrying capacities for females ($K_{f}$) and males ($K_{m}$) were both set to 100 because previous modelling has shown that populations of this size are small enough that mate encounters between kin occur with sufficient frequency for selection on inbreeding strategy, but populations are also not so small that selection is typically overwhelmed by genetic drift (Duthie and Reid *in press*).

@@

We present our results by showing mean allele value changes over generations. Figures 1-5 show mean allele values of tendency for polyandry ($P_{a}$; red lines and shading), inbreeding mating strategy ($M_{a}$; blue lines and shading), and inbreeding fertilisation strategy ($F_{a}$; black lines and shading) over 10000 generations as calculated across all individuals (female and male) within a focal population over 40 replicate simulations with identical starting conditions. Solid lines represent the grand mean of replicate mean allele values, and shading around lines shows standard errors of mean allele values around the grand mean. Panels within figures show different cost combinations for tendency for polyandry ($c_{P}$), inbreeding mating strategy ($c_{M}$), and inbreeding fertilisation strategy ($c_{F}$). Figures have fewer than eight panels if some costs are not relevant because allele values are fixed, and fixed allele values are shown as horizontal lines where line height is set to the fixed allele value. Supporting Information S1 presents correlations of mean allele values across replicate simulations, which were generally weak; we found no evidence suggesting strong evolutionary divergence among replicates with identical starting conditions (e.g., as might be the case if populations stochastically evolved toward inbreeding avoidance through mate choice or fertilisation, but not both), meaning that inspecting grand means of replicate allele values accurately reflects expected evolutionary dynamics over generations.

@@

Inbred offspring are typically less viable than outbred offspring, and when such inbreeding depression is strong, it is predicted to cause selection for inbreeding avoidance [Duthie and Reid *in press*; @Parker1979; @Parker2006; @Pusey1996; @Kokko2006]. In animals, inbreeding avoidance can be realised through both pre-copulatory and post-copulatory mechanisms [e.g., @Tuni2013; @Firman2015; @Daniel2015; @Fischer2015]. However, the conditions under which inbreeding avoidance is predicted to evolve via pre-copulatory, post-copulatory, or both pre- and post-copulatory mechanisms has not been modelled. There is therefore no framework for predicting what mechanisms of inbreeding avoidance are most likely to evolve in empirical populations where risk of fitness loss caused by inbreeding is high. Importantly, evolution of post-copulatory mechanisms of inbreeding avoidance requires females to mate multiply, meaning that post-copulatory inbreeding avoidance cannot occur unless polyandry has evolved in a population. Accordingly, we modelled evolution of three separate traits affecting polyandry and inbreeding avoidance, including tendency to engage in polyandry, inbreeding mating strategy (pre-copulatory), and inbreeding fertilisation strategy (post-copulatory). Using *in silico* experiments, we tested four specific hypothesis to predict evolution of tendency for polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy given diverse costs and evolutionary constraints. We found that (1) selection for polyandry was greatly strengthened given concurrent evolution of inbreeding fertilisation strategy, (2) selection was strong for low cost, and weak but not negligible for high cost, mechanisms of inbreeding avoidance, (3) constraints on evolution of pre-copulatory inbreeding avoidance greatly strengthened selection for post-copulatory inbreeding avoidance, but not vice versa, and (4) pre-existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of post-copulatory inbreeding avoidance, but not vice versa. More generally, we provide a novel framework for approaching empirical studies of inbreeding avoidance, predicting evolution of polyandry and inbreeding avoidance through pre- and post-copulatory mechanisms.

@@

The opportunity to adjust inbreeding has been widely considered as a driver of adaptive evolution of polyandry [@Tregenza2002; @Foerster2003; @Akcay2007; @Varian-Ramos2012; @Kingma2013; @Lehtonen2015; @Reid2014]. Recently, @Duthie found the conditions under which selection for polyandry to avoid inbreeding is predicted to be highly restrictive, requiring strong constraints on the availability of initial compared to additional mates, or the conditional expression of polyandry such that females only engaged in polyandry when their initial choice of mate was of low quality. However, @Duthie assumed that fertilisation of offspring was unbiased with respect to mate kinship, meaning that females could not express choice after mating had occurred. In contrast, we found that when females could express choice after mating through biased fertilisation, selection for polyandry was greatly strengthened (Figure 1A), especially when constraints restricted evolution of inbreeding avoidance through mate choice (Figure 3). The proposition that polyandry might facilitate cryptic female choice among males of varying compatibility is not new [e.g., @Zeh1997], but the existence of selection for post-copulatory inbreeding avoidance causing feedback to select for polyandry has not been formally modelled. Our model therefore has widespread implications for empirical studies of evolution of polyandry.

We predict evolution of polyandry in populations where inbreeding depression is severe and evolution of inbreeding avoidance through post-copulatory mechanisms is not constrained. Post-copulatory mechanisms of inbreeding avoidance have been observed under these conditions in experimental systems across diverse taxa [e.g. @Pizzari2004; @Firman2008; @Bretman2009; @Gasparini2011; @Tuni2013; @Firman2015]. For example, @Pizzari2004 found that the number of sperm arriving to eggs is lower for full siblings of female red jungefowl (*Gallus gallus*) than it is for non-relatives, even after controlling for effects of order of oviposition and social familiarity. Similarly, Firman and Simmons [-@Firman2008; -@Firman2015] have shown evidence of post-copulatory inbreeding avoidance in house mice, finding fertilisation to be biased toward unrelated males independent of the order in which mating occurs [@Firman2008], perhaps due to post-copulatory mechanisms expressed through the secretion of gametic proteins originating in female's eggs [@Firman2015]. Interestingly, female house mice have also been shown to express pre-copulatory inbreeding avoidance  [@Potts1991; @Roberts2003]. Our results suggest that post-copulatory inbreeding avoidance is unlikely to evolve when females can already avoid inbreeding through pre-copulatory mechanisms (Figure 5), meaning that evolution of post-copulatory inbreeding avoidance could reasonably be hypothesised to have preceded evolution of pre-copulatory inbreeding avoidance in mice.

Specific mechanisms of post-copulatory inbreeding avoidance are difficult to identify, but recent experimental studies have shown how females can successfully bias fetilisation after mating has occurred in both vertebrates [e.g., @Gasparini2011] and invertebrates [e.g., @Tuni2013]. For example, @Gasparini2011 artificially inseminated female guppies (*Poecilia reticulata*) with equal quantities of sperm from full-siblings and unrelated males, finding that fertilisation success was higher in non-relatives because the velocity of full siblings' sperm was reduced by females' ovarian fluids. Biased fertilisation toward non-relatives might also be accomplished through selective storage of sperm. For example, in black field crickets (*Teleogryllus commodus*), females attempt to remove the spermataphores of unwanted male mates after copulation, while male mates guard females to prevent spermatophore removal [@Bussiere2006]. Even if females are unable to remove spermatophores of unwanted mates, @Tuni2013 found evidence that female *T. commodus* are capable of controlling the transfer of sperm to their spermatheca, most likely through physical control of the spermathical duct, and can reduce sperm storage despite being guarded by a related male. Our results suggest that evolution of such post-copulatory inbreeding avoidance might increase selection for polyandry, especially if pre-precopulatory inbreeding avoidance is costly (Figure 1) or constrained (Figure 3).

@@

Our results show that direct negative selection on phenotypes that increase female promiscuity or choosiness with respect to kinship decrease the strength of selection for tendency for polyandry and inbreeding avoidance (Figure 1). Importantly, the cost of tendency for polyandry only decreased female fitness given tendency for polyandry phenotype values greater than zero ($P_{p}>0$), but the threshold nature of the tendency for polyandry phenotype only resulted in selection against positive tendency for polyandry allele values ($P_{a}$) when the sum over a focal female's $20$ $P_{a}$ alleles was greater than zero because summations less than zero all resulted in equivalent monandrous phenotypes ($P_{p}=0$). Hence, even when tendency for polyandry was costly, selection for more negative $P_{a}$ values weakened as the frequency of negative $P_{a}$ values increased because each individual $P_{a}$ then became less influential in causing females to be monandrous [@Duthie]. This general property of threshold traits causes maladaptive phenotypes to frequently reappear within populations due to the mutation and recombination of continuous genetic values underlying discrete phenotypes [@Roff1996; @Roff1998]. Assuming that there is continuous genetic variation underlying threshold polyandry, polyandrous phenotypes are therefore sometimes predicted to persist within populations even when costly [@Duthie], potentially creating the possibility for selection on inbreeding fertilisation strategy when polyandry is expressed. This is most evident in Figure 1C, where mean allele values for polyandry are negative (red line), but strong post-copulatory inbreeding avoidance is still predicted (grey line). Consequently, costly polyandry does not necessarily preclude evolution of adaptive post-copulatory inbreeding avoidance in populations.

A low relative cost for one inbreeding strategy (e.g., pre-copulatory inbreeding avoidance) did not preclude a more costly complementary inbreeding strategy from persisting in a focal population (Figure 1), meaning that evolution of both pre- and post-copulatory inbreeding avoidance might be expected in populations despite cost asymmetry. However, existence of costs strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003; @Reid2015]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which we suggest are also relevant costs for polyandry, and include costs of elevated risks associated with (1) predation or (2) disease transmission, and costs incurred through loss of time (3) or depletion of energy (4). In our model, we interpreted a cost of tendency for polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of inbreeding mating strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of inbreeding fertilisation strategy as depletion of energy. While these costs are biologically realistic, different costs will be relevant for different populations and be realised in different ways. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Future models will therefore need to carefully consider the magnitudes and realisation of direct costs to accurately predict evolution of inbreeding strategies in populations. <!--- add Tuni citation for depletion of energy? --->

@@

<!--- BARRIERS to having both pre and post-copulatory inbreeding avoidance evolve are few, although -- all else being equal -- post-copulatory avoidance might evolve less often if there is existing mating avoidance. --->

@@

![Mean allele values underlying tendency for polyandry (red lines and shading), inbreeding mating strategy (i.e., pre-copulatory choice; blue lines and shading), and inbreeding fertilisation strategy (i.e., post-copulatory choice; black lines and shading), as calculated across all individuals within a population over 10000 generations given strong inbreeding depression. Solid lines represent grand means over 40 replicate populations, and shading around solid lines show standard errors of allele values around grand means. Negative mean allele values reflect strategies of inbreeding avoidance or tendency for monandry, and positive values reflect strategies of inbreeding preference or tendency for polyandry. Panels A-H show simulations for different combinations of costs for tendency for polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy. Dotted horizontal black lines in panels show values of zero on the y-axis.](Initial_Results/figures/evo_combs02_B3.png)

![Mean allele values underlying tendency for polyandry (red lines and shading) and inbreeding mating strategy (i.e., pre-copulatory choice; blue lines and shading), as calculated across all individuals within a population over 10000 generations given strong inbreeding depression when allele values underlying inbreeding fertilisation strategy (i.e., post-copulatory choice) are fixed for random fertilisation among mates (black horizontal lines where y-axes equal zero). Solid lines represent grand means over 40 replicate populations, and shading around solid lines show standard errors of allele values around grand means. Negative mean allele values reflect strategies of inbreeding avoidance or tendency for monandry, and positive values reflect strategies of inbreeding preference or tendency for polyandry. Panels A-D show simulations for different combinations of costs for tendency for polyandry and inbreeding mating strategy.](Initial_Results/figures/evo_combs02_B3_ko_F.png)

![Mean allele values underlying tendency for polyandry (red lines and shading) and inbreeding fertilisation strategy (i.e., post-copulatory choice; black lines and shading), as calculated across all individuals within a population over 10000 generations given strong inbreeding depression when allele values underlying inbreeding mating strategy (i.e., pre-copulatory choice) are fixed for random mating (blue horizontal lines where y-axes equal zero). Solid lines represent grand means over 40 replicate populations, and shading around solid lines show standard errors of allele values around grand means. Negative mean allele values reflect strategies of inbreeding avoidance or tendency for monandry, and positive values reflect strategies of inbreeding preference or tendency for polyandry. Panels A-D show simulations for different combinations of costs for tendency for polyandry and inbreeding mating strategy.](Initial_Results/figures/evo_combs02_B3_ko_M-0.png)

![Mean allele values underlying inbreeding mating strategy (i.e., pre-copulatory choice; blue lines and shading), as calculated across all individuals within a population over 10000 generations given strong inbreeding depression and the pre-existence of polyandry to avoid inbreeding (tendency for polyandry allele values equal 1, as indicated by horizontal red lines; inbreeding fertilisation strategy values equal -5, as indicated by horizontal black lines). Solid blue lines represent grand means over 40 replicate populations, and blue shading around solid lines shows standard errors of mean inbreeding mating strategy allele values around grand means. Negative mean allele values reflect a strategy of inbreeding avoidance, and positive values reflect a strategy of inbreeding preference. Panels A and B show replicate simulations in which inbreeding mating strategy is not costly and has a cost of 0.02, respectively. Dotted horizontal black lines in panels show values of zero on the y-axis.](Initial_Results/figures/evo_combs02_B3_ko_FP-5.png)

![Mean allele values underlying tendency for polyandry (red lines and shading) and inbreeding fertilisation strategy (i.e., post-copulatory choice; black lines and shading), as calculated across all individuals within a population over 10000 generations given strong inbreeding depression and the pre-existence of inbreeding avoidance in mate choice (inbreeding mating strategy values equal -5, as indicated by horizontal blue lines). Solid lines represent grand means over 40 replicate populations, and shading around solid lines show standard errors of allele values around grand means. Negative mean allele values reflect strategies of inbreeding avoidance or tendency for monandry, and positive values reflect strategies of inbreeding preference or tendency for polyandry. Panels A-D show simulations for different combinations of costs for tendency for polyandry and inbreeding mating strategy. Dotted horizontal black lines in panels show values of zero on they-axis. ](Initial_Results/figures/evo_combs02_B3_ko_M-5.png)
